NeuronUnit: Collaboratively Testing the Validity of Neuroscientific Models

### 2013-11-30 NeurounitFrontiers/Paper.tex
Some updates to text and moved figure 8 up

The end result of these efforts will be tables, like the one shown in Figure \ref{fig:sciunit_overview}, that allow scientists to understand, precisely, the state-of-the-art in their field. We are developing a lightweight portal, \textit{SciDash}, that directs scientists to the canonical test repositories in their field in order to prevent fragmentation and encourage community participation \citep{omar_sciunit_2013}. \textit{SciDash} will also facilitate viewing test results without needing to clone test repositories and run tests locally for exploratory purposes, based on the \textit{iPython notebook} technology mentioned in Sec. \ref{sec:visualization}. We leave details of \textit{SciDash} as future work, focusing this paper on the core testing methodology (Sec. 2) and on the details of the bridge technologies mentioned above.

\subsection{Reference Data for Tests from NeuroElectro}\label{sec:neuroelectro} 
The NeuroElectro project (\url{http://neuroelectro.org}) is an effort to gather all published data on single cell neurophysiology into machine-readable formats \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are reported for 93 cell types, spanning more than 2000 single pieces of published data extracted from tables in published papers. 
We have made it easy to parameterize \textit{NeuronUnit} tests using data extracted from the NeuroElectro API. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The former is illustrated in Figure \ref{fig:neuronunit_example}. 
Associated statistics of that data (e.g mean, standard error, and sample size) are attached and enable judgement of model output according to a chosen scoring mechanism. 
While NeuroElectro alone cannot judge all model aspects, it can serve to validate basic features of many neurophysiology models, such as resting membrane potential, action potential width, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key component for \textit{NeuronUnit} test constuction.

@@

\begin{figure}
\begin{python}
# Interface with neuroelectro.org to search for spike widths of CA1 Pyramidal cells. 
reference_data = neuroelectro.NeuroElectroSummary(
  neuron = {'name':'Hippocampus CA1 pyramidal cell'}, # Neuron type.  
  ephysprop = {'name':'Spike Width'}) # Electrophysiological property name.
reference_data.get_values()  # Get summary data for the above. 
model = CA1PyramidalCellModel(population_name="CG_CML_0") # Initialize the model with some parameters.
test = SpikeWidthTestDynamic( # Initialize the test.    
	reference_data = {'mean':reference_data.mean, 'std':reference_data.std}, # Summary statistics from the reference data
	model_args = {'current':40.0}, # Somatic current injection in pA.  
	comparator = ZComparator), # A comparison class that implements a Z-score.  
result = sciunit.judge(test,model) # (1) Check capabilities, (2) take the test, (3) generate a score and validate it, (4) bind the score to model/test combination. 
result.summarize() # Summarize the result.  
\end{python}
\caption{Working example of testing in \textit{NeuronUnit}}
\label{fig:neuronunit_example}
\vspace{-15px}
\end{figure}

>>>>

The end result of these efforts will be tables, like the one shown in Figure \ref{fig:sciunit_overview}, that allow scientists to understand, precisely, the state-of-the-art in their field. We are developing a lightweight portal, \textit{SciDash}, that directs scientists to the canonical test repositories in their field in order to prevent fragmentation and encourage community participation \citep{omar_sciunit_2013}. \textit{SciDash} will also facilitate viewing test results without needing to clone test repositories and run tests locally for exploratory purposes, based on the \textit{iPython notebook} technology mentioned in Sec. \ref{sec:visualization}. We leave details of \textit{SciDash} as future work, focusing this paper on the core testing methodology (Sec. 2) and on the details of the bridge technologies.

\subsection{Reference Data for Tests from NeuroElectro}\label{sec:neuroelectro} 
\begin{figure}
\begin{python}
# Interface with neuroelectro.org to search for spike widths of CA1 Pyramidal cells. 
reference_data = neuroelectro.NeuroElectroSummary(
  neuron = {'name':'Hippocampus CA1 pyramidal cell'}, # Neuron type.  
  ephysprop = {'name':'Spike Width'}) # Electrophysiological property name.
reference_data.get_values()  # Get summary data for the above. 
model = CA1PyramidalCellModel(population_name="CG_CML_0") # Initialize the model with some parameters.
test = SpikeWidthTestDynamic( # Initialize the test.    
	reference_data = {'mean':reference_data.mean, 'std':reference_data.std}, # Summary statistics from the reference data
	model_args = {'current':40.0}, # Somatic current injection in pA.  
	comparator = ZComparator), # A comparison class that implements a Z-score.  
result = sciunit.judge(test,model) # (1) Check capabilities, (2) take the test, (3) generate a score and validate it, (4) bind the score to model/test combination. 
result.summarize() # Summarize the result.  
\end{python}
\caption{Working example of testing in \textit{NeuronUnit}}
\label{fig:neuronunit_example}
\vspace{-15px}
\end{figure}
The NeuroElectro project (\url{http://neuroelectro.org}) is an effort to make all published data on single cell neurophysiology available in a machine-readable format \citep{tripathy_neuroelectro:_2012}.  
Currently, up to 27 electrophysiological properties are available for 93 cell types, gathered from over 2000 single pieces of published data extracted from tables in published papers. 
We have made it easy to parameterize \textit{NeuronUnit} tests using data extracted from the NeuroElectro API. 
Tests can be based upon data from single journal articles, or from ensembles of articles with a common theme (e.g. about a particular neuron type). 
The former is illustrated in Figure \ref{fig:neuronunit_example}. 
Associated statistics of that data (e.g mean, standard error, and sample size) are attached and enable judgement of model output according to a chosen scoring mechanism. 
While NeuroElectro alone cannot judge all model aspects, it can serve to validate basic features of many neurophysiology models, such as resting membrane potential, action potential width, after-hyperpolarization amplitude, etc. 
As NeuroElectro is the only publicly curated source of such data, it represents a key component for \textit{NeuronUnit} test constuction.

@@
